---
title: "Genomic Prediction Analysis - Stage II of II: Cross-validation Round 1"
author: "wolfemd"
date: "2019-7-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, tidy = T)
```

# Prep. for cross-validation

##  Read and combine dosages

```{r, eval=F}
library(tidyverse); library(magrittr); library(furrr); library(data.table)
options(mc.cores=18)
plan(multiprocess)
snps<-tibble(Chr=1:18) %>%
    mutate(raw=future_pmap(.,function(Chr,...){
        filename<-paste0("/workdir/marnin/nextgenImputation2019/ImputationStageII_71219/chr",Chr,
                         "_ImputationReferencePanel_StageIIpartI_72219.raw") 
        snps<-fread(filename,
                    stringsAsFactors = F) %>% 
            as_tibble
        return(snps) }))

asModelsFit_nr<-readRDS(file="data/nrcri_blupsForCrossVal_72719.rds")
asModelsFit_iita<-readRDS(file="data/iita_blupsForCrossVal_72619.rds")

asModelsFit_nr %<>% 
      dplyr::select(Trait,blups) %>% 
      unnest(blups) %>% 
      left_join(
            asModelsFit_nr %>% 
                  dplyr::select(Trait,blups) %>% 
                  unnest(blups) %>% 
                  select(GID) %>% 
                  distinct %>% 
                  left_join(tibble(FullSampleName=grep("C1b",rownames(snps),ignore.case = T,value = T)) %>% 
                                  separate(FullSampleName,c("GID","GBS_ID"),":",extra = 'merge',remove = F) %>% 
                                  mutate(GID=paste0("NR16",GID)) %>% 
                                  dplyr::select(-GBS_ID)) %>%
                  mutate(FullSampleName=ifelse(is.na(FullSampleName),GID,FullSampleName)))
asModelsFit_nr %<>% 
      mutate(GID=FullSampleName) %>% 
      dplyr::select(-FullSampleName) %>% 
      group_by(Trait) %>% 
      nest(.key = "blups")

clonesWithBLUPs_nr<-asModelsFit_nr %>% unnest(blups) %$% unique(GID)
clonesWithBLUPs_iita<-asModelsFit_iita %>% unnest(blups) %$% unique(GID)

```
```{r, eval=F}
snps %<>%
  mutate(raw=map(raw,function(raw){
    out<-raw %>% 
      as.data.frame %>% 
      column_to_rownames(var = "IID") %>% 
      dplyr::select(-FID,-PAT,-MAT,-SEX,-PHENOTYPE) %>% 
      as.matrix 
    return(out) }))
snps<-reduce(snps$raw,cbind)

snps_nr<-snps[rownames(snps) %in% clonesWithBLUPs_nr,]
snps_all<-snps[rownames(snps) %in% union(clonesWithBLUPs_iita,
                                         clonesWithBLUPs_nr),]
snps_nrPlusGG<-snps[rownames(snps) %in% grep("TMS13F|TMS14F|TMS15F|2013_",
                                             union(clonesWithBLUPs_iita,clonesWithBLUPs_nr),
                                             value = T,invert = T),]
dim(snps_nr) # [1]  1870 84957
dim(snps_all) # [1]  8132 84957
dim(snps_nrPlusGG) # [1]  2731 84957
```
```{r, eval=F}
asModelsFit_nr %<>%
      mutate(blups=map(blups,~filter(.,GID %in% rownames(snps))))
asModelsFit_nr %>% 
      mutate(Ntrain=map_dbl(blups,nrow)) %>% 
      select(Trait,H2,Ntrain) %>% as.data.frame

```
   Trait           H2 Ntrain
1         DM 0.0252864279    970
2    logRTNO 0.0579231135    972
3         HI 0.1460550481    972
4    logFYLD 0.0918241317    972
5  logTOPYLD 0.1378653091    973
6      MCMDS 0.3339147816   1132
7      CGMS1 0.1513545399   1187
8      CGMS2 0.0396305451   1018
9     BRNHT1 0.0721262328    573
10     PLTHT 0.0008564113    646

## MAF>1% filter
```{r, eval=F}
maf_filter<-function(snps,thresh){
    freq<-colMeans(snps, na.rm=T)/2; maf<-freq;
    maf[which(maf > 0.5)]<-1-maf[which(maf > 0.5)]
    snps1<-snps[,which(maf>thresh)];
    return(snps1) }
snps_nr %<>% maf_filter(.,0.01)
snps_all %<>% maf_filter(.,0.01)
snps_nrPlusGG %<>% maf_filter(.,0.01)
dim(snps_nr) # [1]  1870 83234
dim(snps_all) # [1]  8132 83554
dim(snps_nrPlusGG) # [1]  2731 84150
```

## Make kinships
```{r, eval=F}
library(tidyverse); library(magrittr); library(rrBLUP)
Knr<-rrBLUP::A.mat(snps_nr-1)
KnrPlusGG<-rrBLUP::A.mat(snps_nrPlusGG-1)
Kall<-rrBLUP::A.mat(snps_all-1)

save(Knr,KnrPlusGG,Kall,file=paste0("/workdir/marnin/NRCRI_2019GS/",
                                    "Kinships_NRCRI_ForCrossVal_72819.Rdata"))
saveRDS(snps_nr,file=paste0("/workdir/marnin/NRCRI_2019GS/DosageMatrix_NRCRI_TrainingPop_72819.rds"))
saveRDS(snps_nrPlusGG,file=paste0("/workdir/marnin/NRCRI_2019GS/DosageMatrix_NRCRI_TrainingPop_plusGG_72819.rds"))
saveRDS(snps_all,file=paste0("/workdir/marnin/NRCRI_2019GS/DosageMatrix_NRCRI_TrainingPop_plusAllIITA_72819.rds"))

rm(snps_nr,snps_all,snps_nrPlusGG); gc()
```

# Cross-validation

## Data

```{r, eval=F}
rm(list=ls()); gc()
library(tidyverse); library(magrittr); 
load(file="/workdir/marnin/NRCRI_2019GS/Kinships_NRCRI_ForCrossVal_72819.Rdata")
asModelsFit_nr<-readRDS(file="data/nrcri_blupsForCrossVal_72719.rds")
asModelsFit_nr %<>% 
      dplyr::select(Trait,blups) %>% 
      unnest(blups) %>% 
      left_join(
            asModelsFit_nr %>% 
                  dplyr::select(Trait,blups) %>% 
                  unnest(blups) %>% 
                  select(GID) %>% 
                  distinct %>% 
                  left_join(tibble(FullSampleName=grep("C1b",rownames(Kall),ignore.case = T,value = T)) %>% 
                                  separate(FullSampleName,c("GID","GBS_ID"),":",extra = 'merge',remove = F) %>% 
                                  mutate(GID=paste0("NR16",GID)) %>% 
                                  dplyr::select(-GBS_ID)) %>%
                  mutate(FullSampleName=ifelse(is.na(FullSampleName),GID,FullSampleName)))
asModelsFit_nr %<>% 
      mutate(GID=FullSampleName) %>% 
      dplyr::select(-FullSampleName) %>% 
      group_by(Trait) %>% 
      nest(.key = "blups")

blups<-asModelsFit_nr %>%      
      mutate(blups=map(blups,~filter(.,GID %in% rownames(Kall)))) %>% 
      dplyr::select(Trait,blups) %>% 
      rename(nrTrainingData=blups) %>% 
      left_join(
            readRDS(file="/workdir/marnin/IITA_2019GS/iita_blupsForCrossVal_72619.rds") %>% 
                  rename(trainingData=blups) %>% 
                  mutate(trainingData=map(trainingData,~filter(.,GID %in% rownames(Kall)))) %>% 
                  filter(Trait %in% asModelsFit_nr$Trait) %>% 
                  mutate(keep=ifelse(Dataset=="HistoricalDataIncluded" | Trait %in% c("BRNHT1","PLTHT"),
                                     TRUE,FALSE)) %>% 
                  filter(keep==TRUE) %>% 
                  dplyr::select(Trait,trainingData) %>% 
                  rename(iiTrainingData=trainingData))

c1a<-rownames(Knr) %>% grep("^F",.,value = T); length(c1a) # 899
c1b<-rownames(Knr) %>% grep("C1b",.,value = T,ignore.case = T); length(c1b) # 629
c2<-rownames(Knr) %>% grep("C2a|C2b",.,value = T,ignore.case = T); length(c2) # 57
nrTP<-rownames(Knr) %>% .[!. %in% c(c1a,c1b,c2)]; length(nrTP); # 914
ggEtc<-rownames(Kall) %>% 
      grep("2013_|TMS13F|TMS14F|TMS15F",.,value = T,invert = T) %>% 
      .[!. %in% c(nrTP,c1a,c1b,c2)]; length(ggEtc) # 861
allIITA<-rownames(Kall) %>% .[!. %in% c(nrTP,c1a,c1b,c2)]; length(allIITA) # 6262

blups %<>%
    mutate(seed_of_seeds=1:n(),
           seeds=map(seed_of_seeds,function(seed_of_seeds,reps=5){ 
               set.seed(seed_of_seeds); 
               outSeeds<-sample(1:1000,size = reps,replace = F); 
               return(outSeeds) }))

```
```{bash, eval=F}
mkdir /workdir/NRCRI_2019GS/CrossVal_72819
```

## Function: NRCRI alone

```{r, eval=F}
# nrTrainingData<-blups$nrTrainingData[[1]]; 
# iiTrainingData<-blups$iiTrainingData[[1]];
# seeds<-blups$seeds[[1]]; nfolds<-5; reps<-5;
crossValidateFunc<-function(Trait,nrTrainingData,iiTrainingData,seeds,nfolds=5,reps=5,ncores=40,...){
      K<-Knr
      trntstdata<-nrTrainingData %>% 
            filter(GID %in% rownames(K))
      K1<-K[rownames(K) %in% trntstdata$GID,
            rownames(K) %in% trntstdata$GID]
      rm(Knr,KnrPlusGG,Kall,K,nrTrainingData,iiTrainingData); gc()
      # seed<-seeds[[1]]
      # Nfolds=nfolds
      makeFolds<-function(Nfolds=nfolds,seed){
            genotypes<-rownames(K1)
            
            set.seed(seed)
            seed_per_group<-sample(1:10000,size = 3,replace = FALSE)
      
            set.seed(seed_per_group[1])
            FoldsThisRep_nrTP<-tibble(CLONE=genotypes[genotypes %in% nrTP],
                                       Group="nrTP") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[2])
            FoldsThisRep_c1a<-tibble(CLONE=genotypes[genotypes %in% c1a],
                                       Group="c1a") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[3])
            FoldsThisRep_c1b<-tibble(CLONE=genotypes[genotypes %in% c1b],
                                       Group="c1b") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
           FoldsThisRep<-bind_rows(FoldsThisRep_nrTP,FoldsThisRep_c1a) %>% 
                  bind_rows(FoldsThisRep_c1b) %>%  
                  mutate(Test=map(Test,~.$CLONE),
                         Train=map(Test,~genotypes[!genotypes %in% .]))
            return(FoldsThisRep) }
      
    crossval<-tibble(Rep=1:reps,seed=unlist(seeds)) %>% 
        mutate(Folds=map2(Rep,seed,~makeFolds(Nfolds=nfolds,seed=.y))) %>% 
        unnest()

    #Test<-crossval$Test[[1]]; Train<-crossval$Train[[1]]
    crossValidate<-function(Test,Train){ 
      train<-Train
      test<-Test
      trainingdata<-trntstdata %>% 
        filter(GID %in% train) %>% 
        mutate(GID=factor(GID,levels=rownames(K1)))
       
      require(sommer)
      proctime<-proc.time()
      fit <- mmer(fixed = drgBLUP ~1,
                  random = ~vs(GID,Gu=K1),
                  weights = WT,
                  data=trainingdata) 
      proc.time()-proctime
      
      x<-fit$U$`u:GID`$drgBLUP 
      gebvs<-tibble(GID=names(x),
                    GEBV=as.numeric(x))
      
      accuracy<-gebvs %>% 
            filter(GID %in% test) %>% 
            left_join(
                  trntstdata %>% 
                        dplyr::select(GID,BLUP) %>% 
                        filter(GID %in% test)) %$% 
            cor(GEBV,BLUP, use='complete.obs') 
    return(accuracy)
    }
    
    require(furrr)
    options(mc.cores=ncores)
    plan(multiprocess)
    crossval<-crossval %>% 
          mutate(accuracy=future_map2(Test,Train,~crossValidate(Test=.x,Train=.y)))
    saveRDS(crossval,file=paste0("output/CrossVal_72819/",
                                 "CrossVal_",Trait,"_NRCRIonly_72819.rds"))
    rm(list=ls()); gc() 
    }
```

### cbsumm21 (40)

```{r, eval=F}
blups %>% 
      slice(1:5) %>% 
      mutate(CVaccuracy=pmap(.,crossValidateFunc))
```

### cbsu30 (40)

```{r, eval=F}
blups %>% 
      slice(6:10) %>% 
      mutate(CVaccuracy=pmap(.,crossValidateFunc))
```

## Function: IITA augmented

```{r, eval=F}
# nrTrainingData<-blups$nrTrainingData[[1]]; 
# iiTrainingData<-blups$iiTrainingData[[1]];
# seeds<-blups$seeds[[1]]; nfolds<-5; reps<-5;
crossValidateFunc<-function(Trait,nrTrainingData,iiTrainingData,seeds,nfolds=5,reps=5,ncores=30,...){
      K<-Kall
      trntstdata<-nrTrainingData %>% 
            mutate(Dataset="NRCRI") %>% 
            bind_rows(iiTrainingData %>% 
                            mutate(Dataset="IITA")) %>% 
            filter(GID %in% rownames(K))
      K1<-K[rownames(K) %in% trntstdata$GID,
            rownames(K) %in% trntstdata$GID]
      rm(Knr,KnrPlusGG,Kall,K,nrTrainingData,iiTrainingData); gc()
      # seed<-seeds[[1]]
      # Nfolds=nfolds
      makeFolds<-function(Nfolds=nfolds,seed){
            genotypes<-rownames(K1)
            
            set.seed(seed)
            seed_per_group<-sample(1:10000,size = 3,replace = FALSE)
      
            set.seed(seed_per_group[1])
            FoldsThisRep_nrTP<-tibble(CLONE=genotypes[genotypes %in% nrTP],
                                       Group="nrTP") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[2])
            FoldsThisRep_c1a<-tibble(CLONE=genotypes[genotypes %in% c1a],
                                       Group="c1a") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[3])
            FoldsThisRep_c1b<-tibble(CLONE=genotypes[genotypes %in% c1b],
                                       Group="c1b") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
           FoldsThisRep<-bind_rows(FoldsThisRep_nrTP,FoldsThisRep_c1a) %>% 
                  bind_rows(FoldsThisRep_c1b) %>%  
                  mutate(Test=map(Test,~.$CLONE),
                         Train=map(Test,~genotypes[!genotypes %in% .]))
            return(FoldsThisRep) }
      
    crossval<-tibble(Rep=1:reps,seed=unlist(seeds)) %>% 
        mutate(Folds=map2(Rep,seed,~makeFolds(Nfolds=nfolds,seed=.y))) %>% 
        unnest()
    #Test<-crossval$Test[[1]]; Train<-crossval$Train[[1]]
    crossValidate<-function(Test,Train){ 
      train<-Train
      test<-Test
      trainingdata<-trntstdata %>% 
        filter(GID %in% train) %>% 
        mutate(GID=factor(GID,levels=rownames(K1)))
       
      require(sommer)
      proctime<-proc.time()
      fit <- mmer(fixed = drgBLUP ~1,
                  random = ~vs(GID,Gu=K1),
                  weights = WT,
                  data=trainingdata) 
      proc.time()-proctime
      
      x<-fit$U$`u:GID`$drgBLUP 
      gebvs<-tibble(GID=names(x),
                    GEBV=as.numeric(x))
      
      accuracy<-gebvs %>% 
            filter(GID %in% test) %>% 
            left_join(
                  trntstdata %>% 
                        filter(Dataset=="NRCRI") %>% # only validate on NRCRI blups
                        dplyr::select(GID,BLUP) %>% 
                        filter(GID %in% test)) %$% 
            cor(GEBV,BLUP, use='complete.obs') 
    return(accuracy)
    }
    
    require(furrr)
    options(mc.cores=ncores)
    plan(multiprocess)
    crossval<-crossval %>% 
          mutate(accuracy=future_map2(Test,Train,~crossValidate(Test=.x,Train=.y)))
    saveRDS(crossval,file=paste0("output/CrossVal_72819/",
                                 "CrossVal_",Trait,"_IITAaugmented_72819.rds"))
    rm(list=ls()); gc() 
    }
```

### cbsumm21 (40)

```{r, eval=F}
blups %>% 
      slice(1:5) %>% 
      mutate(CVaccuracy=pmap(.,crossValidateFunc))
```

### cbsu30 (40)

```{r, eval=F}
blups %>% 
      slice(6:10) %>% 
      mutate(CVaccuracy=pmap(.,crossValidateFunc))
```

## Plot Results

```{r, eval=T}
rm(list=ls());gc()
library(tidyverse); library(magrittr); library(cowplot); 
cv<-tibble(Files=list.files("output/CrossVal_72819/")) %>% 
      mutate(cvResults=map(Files,~readRDS(paste0("output/CrossVal_72819/",.))))
cv %<>% 
      mutate(Files=gsub("CrossVal_","",Files),
             Files=gsub("_72819.rds","",Files)) %>% 
      separate(Files,c("Trait","Dataset"),"_")
cv %<>% 
  unnest(cvResults)
cv %<>% 
  dplyr::select(-Test,-Train) %>% 
  unnest(accuracy)
```

```{r, fig.width=9, fig.height=6}
library(viridis)
cv %>% 
      mutate(Group=factor(Group,levels=c("nrTP","c1a","c1b")),
             Dataset=factor(Dataset,levels=c("NRCRIonly","IITAaugmented"))) %>% 
      ggplot(.,aes(x=Trait,y=accuracy,fill=Dataset)) + 
      geom_boxplot() + 
      facet_grid(.~Group,space = 'free', scales = 'free_x') + 
      geom_hline(yintercept = 0,color='darkred',size=1.25) + 
      theme_bw() + 
      theme(axis.text.x = element_text(angle=90)) +
      scale_fill_viridis_d()
```

# Cross-validation on curated data

## Prep. data for cross-val

```{r, eval=F}
rm(list=ls()); gc()
library(tidyverse); library(magrittr); 
blups<-readRDS(file="data/nrcri_blupsForCrossVal_outliersRemoved_73019.rds")
K<-readRDS(file=paste0("/workdir/marnin/NRCRI_2019GS/Kinship_IITA_TrainingPop_72619.rds"))
blups %<>%
      rename(trainingData=blups) %>% 
      mutate(trainingData=map(trainingData,~filter(.,GID %in% rownames(K))),)
tms13f<-rownames(K) %>% grep("TMS13F|2013_",.,value = T); length(tms13f) # 2395
tms14f<-rownames(K) %>% grep("TMS14F",.,value = T); length(tms14f) # 2171
tms15f<-rownames(K) %>% grep("TMS15F",.,value = T); length(tms15f) # 835
gg<-setdiff(rownames(K),c(tms13f,tms14f,tms15f)); length(gg) # 1228 (not strictly gg)

blups %<>%
    mutate(seed_of_seeds=1:n(),
           seeds=map(seed_of_seeds,function(seed_of_seeds,reps=5){ 
               set.seed(seed_of_seeds); 
               outSeeds<-sample(1:1000,size = reps,replace = F); 
               return(outSeeds) }))
blups %<>%
      select(-varcomp); gc()
```

```{r, eval=F}
rm(list=ls()); gc()
library(tidyverse); library(magrittr); 
load(file="/workdir/marnin/NRCRI_2019GS/Kinships_NRCRI_ForCrossVal_72819.Rdata")
asModelsFit_nr<-readRDS(file="data/nrcri_blupsForCrossVal_outliersRemoved_73019.rds")
asModelsFit_nr %<>% 
      dplyr::select(Trait,blups) %>% 
      unnest(blups) %>% 
      left_join(
            asModelsFit_nr %>% 
                  dplyr::select(Trait,blups) %>% 
                  unnest(blups) %>% 
                  select(GID) %>% 
                  distinct %>% 
                  left_join(tibble(FullSampleName=grep("C1b",rownames(Kall),ignore.case = T,value = T)) %>% 
                                  separate(FullSampleName,c("GID","GBS_ID"),":",extra = 'merge',remove = F) %>% 
                                  mutate(GID=paste0("NR16",GID)) %>% 
                                  dplyr::select(-GBS_ID)) %>%
                  mutate(FullSampleName=ifelse(is.na(FullSampleName),GID,FullSampleName)))
asModelsFit_nr %<>% 
      mutate(GID=FullSampleName) %>% 
      dplyr::select(-FullSampleName) %>% 
      group_by(Trait) %>% 
      nest(.key = "blups")

blups<-asModelsFit_nr %>%      
      mutate(blups=map(blups,~filter(.,GID %in% rownames(Kall)))) %>% 
      dplyr::select(Trait,blups) %>% 
      rename(nrTrainingData=blups) %>% 
      left_join(
            readRDS(file="data/iita_blupsForCrossVal_outliersRemoved_73019.rds") %>% 
                  rename(trainingData=blups) %>% 
                  mutate(trainingData=map(trainingData,~filter(.,GID %in% rownames(Kall)))) %>% 
                  filter(Trait %in% asModelsFit_nr$Trait) %>% 
                  dplyr::select(Trait,trainingData) %>% 
                  rename(iiTrainingData=trainingData))

c1a<-rownames(Knr) %>% grep("^F",.,value = T); length(c1a) # 899
c1b<-rownames(Knr) %>% grep("C1b",.,value = T,ignore.case = T); length(c1b) # 629
c2<-rownames(Knr) %>% grep("C2a|C2b",.,value = T,ignore.case = T); length(c2) # 57
nrTP<-rownames(Knr) %>% .[!. %in% c(c1a,c1b,c2)]; length(nrTP); # 914
ggEtc<-rownames(Kall) %>% 
      grep("2013_|TMS13F|TMS14F|TMS15F",.,value = T,invert = T) %>% 
      .[!. %in% c(nrTP,c1a,c1b,c2)]; length(ggEtc) # 861
allIITA<-rownames(Kall) %>% .[!. %in% c(nrTP,c1a,c1b,c2)]; length(allIITA) # 6262

blups %<>%
    mutate(seed_of_seeds=1:n(),
           seeds=map(seed_of_seeds,function(seed_of_seeds,reps=5){ 
               set.seed(seed_of_seeds); 
               outSeeds<-sample(1:1000,size = reps,replace = F); 
               return(outSeeds) }))

```
```{bash, eval=F}
mkdir /workdir/NRCRI_2019GS/CrossVal_73019
```

## NRCRI alone
```{r, eval=F}
# nrTrainingData<-blups$nrTrainingData[[1]]; 
# iiTrainingData<-blups$iiTrainingData[[1]];
# seeds<-blups$seeds[[1]]; nfolds<-5; reps<-5;

crossValidateFunc<-function(Trait,nrTrainingData,iiTrainingData,seeds,nfolds=5,reps=5,ncores=100,...){
      K<-Knr
      trntstdata<-nrTrainingData %>% 
            filter(GID %in% rownames(K))
      K1<-K[rownames(K) %in% trntstdata$GID,
            rownames(K) %in% trntstdata$GID]
      rm(Knr,KnrPlusGG,Kall,K,nrTrainingData,iiTrainingData); gc()
      # seed<-seeds[[1]]
      # Nfolds=nfolds
      makeFolds<-function(Nfolds=nfolds,seed){
            genotypes<-rownames(K1)
            
            set.seed(seed)
            seed_per_group<-sample(1:10000,size = 3,replace = FALSE)
      
            set.seed(seed_per_group[1])
            FoldsThisRep_nrTP<-tibble(CLONE=genotypes[genotypes %in% nrTP],
                                       Group="nrTP") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[2])
            FoldsThisRep_c1a<-tibble(CLONE=genotypes[genotypes %in% c1a],
                                       Group="c1a") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[3])
            FoldsThisRep_c1b<-tibble(CLONE=genotypes[genotypes %in% c1b],
                                       Group="c1b") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
           FoldsThisRep<-bind_rows(FoldsThisRep_nrTP,FoldsThisRep_c1a) %>% 
                  bind_rows(FoldsThisRep_c1b) %>%  
                  mutate(Test=map(Test,~.$CLONE),
                         Train=map(Test,~genotypes[!genotypes %in% .]))
            return(FoldsThisRep) }
      
    crossval<-tibble(Rep=1:reps,seed=unlist(seeds)) %>% 
        mutate(Folds=map2(Rep,seed,~makeFolds(Nfolds=nfolds,seed=.y))) %>% 
        unnest()

    #Test<-crossval$Test[[1]]; Train<-crossval$Train[[1]]
    crossValidate<-function(Test,Train){ 
      train<-Train
      test<-Test
      trainingdata<-trntstdata %>% 
        filter(GID %in% train) %>% 
        mutate(GID=factor(GID,levels=rownames(K1)))
       
      require(sommer)
      proctime<-proc.time()
      fit <- mmer(fixed = drgBLUP ~1,
                  random = ~vs(GID,Gu=K1),
                  weights = WT,
                  data=trainingdata) 
      proc.time()-proctime
      
      x<-fit$U$`u:GID`$drgBLUP 
      gebvs<-tibble(GID=names(x),
                    GEBV=as.numeric(x))
      
      accuracy<-gebvs %>% 
            filter(GID %in% test) %>% 
            left_join(
                  trntstdata %>% 
                        dplyr::select(GID,BLUP) %>% 
                        filter(GID %in% test)) %$% 
            cor(GEBV,BLUP, use='complete.obs') 
    return(accuracy)
    }
    
    require(furrr)
    options(mc.cores=ncores)
    plan(multiprocess)
    crossval<-crossval %>% 
          mutate(accuracy=future_map2(Test,Train,~crossValidate(Test=.x,Train=.y)))
    saveRDS(crossval,file=paste0("output/CrossVal_73019/",
                                 "CrossVal_",Trait,"_NRCRIonly_OutliersRemoved_73019.rds"))
    rm(list=ls()); gc() 
    }
```

### cbsulm16 (112)
```{r, eval=F}
blups %>% 
      mutate(CVaccuracy=pmap(.,crossValidateFunc))
```

## IITA augmented
```{r, eval=F}
# nrTrainingData<-blups$nrTrainingData[[1]]; 
# iiTrainingData<-blups$iiTrainingData[[1]];
# seeds<-blups$seeds[[1]]; nfolds<-5; reps<-5;
crossValidateFunc<-function(Trait,nrTrainingData,iiTrainingData,seeds,nfolds=5,reps=5,ncores=100,...){
      K<-Kall
      trntstdata<-nrTrainingData %>% 
            mutate(Dataset="NRCRI") %>% 
            bind_rows(iiTrainingData %>% 
                            mutate(Dataset="IITA")) %>% 
            filter(GID %in% rownames(K))
      K1<-K[rownames(K) %in% trntstdata$GID,
            rownames(K) %in% trntstdata$GID]
      rm(Knr,KnrPlusGG,Kall,K,nrTrainingData,iiTrainingData); gc()
      # seed<-seeds[[1]]
      # Nfolds=nfolds
      makeFolds<-function(Nfolds=nfolds,seed){
            genotypes<-rownames(K1)
            
            set.seed(seed)
            seed_per_group<-sample(1:10000,size = 3,replace = FALSE)
      
            set.seed(seed_per_group[1])
            FoldsThisRep_nrTP<-tibble(CLONE=genotypes[genotypes %in% nrTP],
                                       Group="nrTP") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[2])
            FoldsThisRep_c1a<-tibble(CLONE=genotypes[genotypes %in% c1a],
                                       Group="c1a") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
            set.seed(seed_per_group[3])
            FoldsThisRep_c1b<-tibble(CLONE=genotypes[genotypes %in% c1b],
                                       Group="c1b") %>% 
                  mutate(Folds=sample(1:Nfolds,nrow(.),replace=T)) %>% 
                  arrange(Folds) %>% 
                  group_by(Group,Folds) %>% 
                  nest(.key = Test)
           FoldsThisRep<-bind_rows(FoldsThisRep_nrTP,FoldsThisRep_c1a) %>% 
                  bind_rows(FoldsThisRep_c1b) %>%  
                  mutate(Test=map(Test,~.$CLONE),
                         Train=map(Test,~genotypes[!genotypes %in% .]))
            return(FoldsThisRep) }
      
      crossval<-tibble(Rep=1:reps,seed=unlist(seeds)[1:reps]) %>% 
        mutate(Folds=map2(Rep,seed,~makeFolds(Nfolds=nfolds,seed=.y))) %>% 
        unnest()
    #Test<-crossval$Test[[1]]; Train<-crossval$Train[[1]]
    crossValidate<-function(Test,Train){ 
      train<-Train
      test<-Test
      trainingdata<-trntstdata %>% 
        filter(GID %in% train) %>% 
        mutate(GID=factor(GID,levels=rownames(K1)))
       
      require(sommer)
      proctime<-proc.time()
      fit <- mmer(fixed = drgBLUP ~1,
                  random = ~vs(GID,Gu=K1),
                  weights = WT,
                  data=trainingdata) 
      proc.time()-proctime
      
      x<-fit$U$`u:GID`$drgBLUP 
      gebvs<-tibble(GID=names(x),
                    GEBV=as.numeric(x))
      
      accuracy<-gebvs %>% 
            filter(GID %in% test) %>% 
            left_join(
                  trntstdata %>% 
                        filter(Dataset=="NRCRI") %>% # only validate on NRCRI blups
                        dplyr::select(GID,BLUP) %>% 
                        filter(GID %in% test)) %$% 
            cor(GEBV,BLUP, use='complete.obs') 
    return(accuracy)
    }
    
    require(furrr)
    options(mc.cores=ncores)
    plan(multiprocess)
    crossval<-crossval %>% 
          mutate(accuracy=future_map2(Test,Train,~crossValidate(Test=.x,Train=.y)))
    saveRDS(crossval,file=paste0("/output/CrossVal_73019/",
                                 "CrossVal_",Trait,"_IITAaugmented_OutliersRemoved_73019.rds"))
    rm(list=ls()); gc() 
    }
```

### cbsulm16 (112)

```{r, eval=F}
blups %>% 
      mutate(CVaccuracy=pmap(.,crossValidateFunc))
```

## Plot Results
```{r, eval=T}
rm(list=ls());gc()
library(tidyverse); library(magrittr); library(cowplot); 
cvNoOutliers<-tibble(Files=list.files("output/CrossVal_73019/")) %>% 
                  mutate(Trait=gsub("CrossVal_","",Files),
                         Trait=gsub("_OutliersRemoved_73019.rds","",Trait),
                         Curation="OutliersRemoved") %>% 
      separate(Trait,c("Trait","Dataset"),"_") %>% 
      mutate(cvResults=map(Files,~readRDS(paste0("output/CrossVal_73019/",.)))) %>% 
      dplyr::select(-Files)


cvWithOutliers<-tibble(Files=list.files("output/CrossVal_72819/")) %>% 
      mutate(Trait=gsub("CrossVal_","",Files),
             Trait=gsub("_72819.rds","",Trait),
             Curation="NoOutlierRemoval") %>% 
      separate(Trait,c("Trait","Dataset"),"_") %>% 
      mutate(cvResults=map(Files,~readRDS(paste0("output/CrossVal_72819/",.)))) %>% 
      dplyr::select(-Files)
cv<-bind_rows(cvNoOutliers,
              cvWithOutliers)

cv %<>%  
  unnest(cvResults) %>% 
  dplyr::select(-Test,-Train) %>% 
  unnest(accuracy)
```


```{r, fig.width=9, fig.height=6}
library(viridis)
cv %>% 
      ggplot(.,aes(x=Trait,y=accuracy,fill=Curation)) + 
      geom_boxplot() + 
      facet_grid(Group~Dataset,space='free_x',scale='free_x') +
      geom_hline(yintercept = 0,color='darkred',size=1.25) + 
      theme_bw() + 
      theme(axis.text.x = element_text(angle=90,face='bold',size=14)) +
      scale_fill_viridis_d()
```

# Next step
[Stage II: Get GEBVs](GetGEBVs.html)
