---
title: "Predict GEBV for NRCRI C2"
author: "wolfemd"
date: "2019-7-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, tidy = T)
```

# Load snps

```{r, eval=F}
rm(list=ls()); gc()
library(tidyverse); library(magrittr); 
snps<-readRDS(file=paste0("/workdir/marnin/nextgenImputation2019/ImputationStageIII_72619/",
                         "DosageMatrix_RefPanelAndGSprogeny_ReadyForGP_73019.rds"))
```

# Load blups

## IITA 

```{r, eval=F}
blupsNoOut<-readRDS(file="data/iita_blupsForCrossVal_outliersRemoved_73019.rds")
blups<-readRDS(file="data/iita_blupsForCrossVal_72619.rds")

blups<-blupsNoOut %>% 
      dplyr::select(Trait,blups) %>% 
      mutate(OutliersRemoved=TRUE) %>% 
      bind_rows(
            blups %>% 
                  filter(Dataset=="HistoricalDataIncluded" | Trait %in% c("BRNHT1","PLTHT")) %>% 
                  dplyr::select(Trait,blups) %>% 
                  mutate(OutliersRemoved=FALSE))
gidsInPhenos<-tibble(GID=blups %>% unnest() %$% GID %>% unique)
gidsInPhenos %<>% 
      left_join(
            tibble(dartName=rownames(snps) %>% grep("TMS18",.,value=T)) %>% 
                  separate(dartName,c("GID","dartID"),"_",extra = 'merge',remove = F) %>% 
                  semi_join(gidsInPhenos) %>% 
                  group_by(GID) %>% 
                  slice(1) %>% 
                  ungroup()) %>% 
      mutate(NewName=ifelse(!is.na(dartName),dartName,GID))
blups %<>% 
      unnest(blups) %>% 
      left_join(gidsInPhenos) %>% 
      mutate(GID=NewName) %>% 
      dplyr::select(-dartName,-dartID,-NewName) %>% 
      filter(GID %in% rownames(snps)) %>% 
      group_by(Trait,OutliersRemoved) %>% 
      nest(.key = "trainingData")

iitablups<-blups
iitaGIDsInPhenos<-gidsInPhenos
```

## NRCRI

```{r, eval=F}
blupsNoOut<-readRDS(file="data/nrcri_blupsForCrossVal_72719.rds")
blups<-readRDS(file="data/nrcri_blupsForCrossVal_outliersRemoved_73019.rds")
blups<-blupsNoOut %>% 
      dplyr::select(Trait,blups) %>% 
      mutate(OutliersRemoved=TRUE) %>% 
      bind_rows(
            blups %>% 
                  dplyr::select(Trait,blups) %>% 
                  mutate(OutliersRemoved=FALSE))
gidsInPhenos<-tibble(GID=blups %>% unnest() %$% GID %>% unique)
table(gidsInPhenos$GID %in% rownames(snps)) # 1241
gidsInPhenos$GID[gidsInPhenos$GID %in% rownames(snps)] %>% grep("C1b",.,ignore.case = T, value = T) # none

gidsInPhenos %<>% 
      mutate(NewName=GID) %>%
      filter(NewName %in% rownames(snps)) %>% 
      bind_rows(
            gidsInPhenos %>% 
                  mutate(NewName=GID) %>%
                  filter(!NewName %in% rownames(snps)) %>% 
                  left_join(
                        tibble(FullSampleName=grep("C1b",rownames(snps),ignore.case = T,value = T)) %>% 
                              separate(FullSampleName,c("NewName","GBS_ID"),":",extra = 'merge',remove = F) %>% 
                              mutate(NewName=paste0("NR16",NewName)) %>% 
                              dplyr::select(-GBS_ID)) %>% 
                  filter(!is.na(FullSampleName)) %>% 
                  mutate(NewName=FullSampleName) %>% 
                  dplyr::select(-FullSampleName)) %>% 
      bind_rows(
            gidsInPhenos %>% 
                  mutate(NewName=GID) %>%
                  filter(!NewName %in% rownames(snps),
                         grepl("C2a",NewName,ignore.case = T)) %>% 
                  left_join(
                        tibble(dartName=grep("C2a",rownames(snps),ignore.case = T,value = T)) %>% 
                              separate(dartName,c("dartID","NewName"),"\\.\\.\\.",extra = 'merge',remove = F) %>% 
                              filter(!is.na(NewName))) %>% 
                  mutate(NewName=ifelse(!is.na(dartName),dartName,NewName)) %>% 
                  dplyr::select(-dartName,-dartID)) %>% 
      filter(NewName %in% rownames(snps))
table(gidsInPhenos$NewName %in% rownames(snps)) # 2768

blups %<>% 
      unnest(blups) %>% 
      left_join(gidsInPhenos) %>% 
      filter(!is.na(NewName)) %>% 
      mutate(GID=NewName) %>% 
      dplyr::select(-NewName) %>% 
      group_by(Trait,OutliersRemoved) %>% 
      nest(.key = "trainingData")
```

# Make a kinship matrix

## Subset snps
```{r, eval=F}
samples2Keep<-union(iitaGIDsInPhenos$NewName %>% .[. %in% rownames(snps)],
                    gidsInPhenos$NewName %>% .[. %in% rownames(snps)]) %>% 
      union(.,rownames(snps) %>% grep("C2",.,value = T, ignore.case = T))

length(samples2Keep) # 13363
snps<-snps[samples2Keep,]
```

## MAF>1% filter
```{r, eval=F}
maf_filter<-function(snps,thresh){
    freq<-colMeans(snps, na.rm=T)/2; maf<-freq;
    maf[which(maf > 0.5)]<-1-maf[which(maf > 0.5)]
    snps1<-snps[,which(maf>thresh)];
    return(snps1) }
snps %<>% maf_filter(.,0.01)
dim(snps) # [1] 13363 68695
```

## Make kinships
```{r, eval=F}
library(tidyverse); library(magrittr); library(rrBLUP)
K<-rrBLUP::A.mat(snps-1)

saveRDS(K,file=paste0("/workdir/marnin/NRCRI_2019GS/",
                      "Kinship_NRCRI_SamplesForGP_73119.rds"))
saveRDS(snps,file=paste0("/workdir/marnin/NRCRI_2019GS/",
                         "DosageMatrix_NRCRI_SamplesForGP_73119.rds"))
#rm(snps); gc()
```

# PCA 
```{r, eval=F}
pca<-prcomp(snps, scale=T, center=T) 
pc_scores<-pca$x %>% 
    .[,1:15] %>% 
    as.data.frame %>% 
    rownames_to_column(var="FullSampleName") %>% 
    mutate(GeneticGroup=NA,
           GeneticGroup=ifelse(FullSampleName %in% iitaGIDsInPhenos$NewName,"IITA",
                               ifelse(grepl("C2aF",FullSampleName,ignore.case = T),"C2a",
                                      ifelse(grepl("C2bF",FullSampleName,ignore.case = T),"C2b",
                                             ifelse(grep("^F",FullSampleName,value = T),"C1a",
                                                    ifelse(grepl("C1b",FullSampleName,ignore.case = T),"C1b","nrTP"))))))
pc_scores %>%
      count(GeneticGroup)
summary(pca)$importance[,1:10] 
```
                            PC1      PC2      PC3      PC4      PC5      PC6
Standard deviation     59.20352 48.62852 42.35600 36.86073 35.90890 33.34435
Proportion of Variance  0.05102  0.03442  0.02612  0.01978  0.01877  0.01619
Cumulative Proportion   0.05102  0.08545  0.11156  0.13134  0.15011  0.16630
                            PC7      PC8      PC9     PC10
Standard deviation     32.97758 30.70189 30.22302 27.44181
Proportion of Variance  0.01583  0.01372  0.01330  0.01096
Cumulative Proportion   0.18213  0.19585  0.20915  0.22011

Save PCA results
```{r, eval=F}
saveRDS(pc_scores,file="/workdir/marnin/NRCRI_2019GS/PCscores_NRCRI_SamplesForGP_73119.rds")
saveRDS(pca,file="/workdir/marnin/NRCRI_2019GS/PCA_NRCRI_SamplesForGP_73119.rds")
rm(pca); gc()
```
Copy to local machine
 0.04623  0.02847  0.02576  0.02201
 
```{r, fig.width=9, fig.height=6}
#rm(list=ls());gc()
library(tidyverse); library(magrittr); library(cowplot); 
pc_scores<-readRDS("output/PCscores_NRCRI_SamplesForGP_73119.rds")
#pc_scores$FullSampleName %>% grep("C1b",.,value = T,ignore.case = T)
pc_scores %<>% 
       mutate(GeneticGroup=NA,
           GeneticGroup=ifelse(grepl("NR17F",FullSampleName,ignore.case = T),"C2a",
                                      ifelse(grepl("C2bF",FullSampleName,ignore.case = T),"C2b",
                                             ifelse(grepl("^F",FullSampleName) & !grepl("C1b",FullSampleName),"C1a",
                                                    ifelse(grepl("C1b",FullSampleName,
                                                                 ignore.case = T),"C1b","nrTP_plus_IITA"))))) 
library(viridis)
pc1v2<-pc_scores %>% 
      ggplot(.,aes(x=PC1,y=PC2,color=GeneticGroup)) + 
      geom_point() + 
      theme_bw() + scale_color_viridis_d() + 
      labs(x="PC1 (5%)",y="PC2 (3%)") + theme(legend.position = 'none')
pc3v4<-pc_scores %>% 
      ggplot(.,aes(x=PC3,y=PC4,color=GeneticGroup)) + 
      geom_point() + 
      theme_bw() + scale_color_viridis_d() + 
      labs(x="PC3 (3%)",y="PC4 (2%)")
plot_grid(pc1v2,pc3v4)
```
```{r, fig.width=9, fig.height=6}
pc_scores %>% 
      ggplot(.,aes(x=PC5,y=PC6,color=GeneticGroup)) + 
      geom_point() + 
      theme_bw() + scale_color_viridis_d() + 
      labs(x="PC5 (2%)",y="PC6 (2%)")
```

# Prediction

## Load data for pred
```{r, eval=F}
rm(list=ls()); gc()
library(tidyverse); library(magrittr); 
K<-readRDS(file=paste0("/workdir/NRCRI_2019GS/Kinship_NRCRI_SamplesForGP_73119.rds"))

blupsNoOut<-readRDS(file="data/iita_blupsForCrossVal_outliersRemoved_73019.rds")
blups<-readRDS(file="data/iita_blupsForCrossVal_72619.rds")
blups<-blupsNoOut %>% 
      dplyr::select(Trait,blups) %>% 
      mutate(OutliersRemoved=TRUE) %>% 
      bind_rows(
            blups %>% 
                  filter(Dataset=="HistoricalDataIncluded" | Trait %in% c("BRNHT1","PLTHT")) %>% 
                  dplyr::select(Trait,blups) %>% 
                  mutate(OutliersRemoved=FALSE))
gidsInPhenos<-tibble(GID=blups %>% unnest() %$% GID %>% unique)
gidsInPhenos %<>% 
      left_join(
            tibble(dartName=rownames(K) %>% grep("TMS18",.,value=T)) %>% 
                  separate(dartName,c("GID","dartID"),"_",extra = 'merge',remove = F) %>% 
                  semi_join(gidsInPhenos) %>% 
                  group_by(GID) %>% 
                  slice(1) %>% 
                  ungroup()) %>% 
      mutate(NewName=ifelse(!is.na(dartName),dartName,GID))
blups %<>% 
      unnest(blups) %>% 
      left_join(gidsInPhenos) %>% 
      mutate(GID=NewName) %>% 
      dplyr::select(-dartName,-dartID,-NewName) %>% 
      filter(GID %in% rownames(K)) %>% 
      group_by(Trait,OutliersRemoved) %>% 
      nest(.key = "trainingData")
iitablups<-blups
iitaGIDsInPhenos<-gidsInPhenos

blupsNoOut<-readRDS(file="data/nrcri_blupsForCrossVal_72719.rds")
blups<-readRDS(file="data/nrcri_blupsForCrossVal_outliersRemoved_73019.rds")
blups<-blupsNoOut %>% 
      dplyr::select(Trait,blups) %>% 
      mutate(OutliersRemoved=TRUE) %>% 
      bind_rows(
            blups %>% 
                  dplyr::select(Trait,blups) %>% 
                  mutate(OutliersRemoved=FALSE))
gidsInPhenos<-tibble(GID=blups %>% unnest() %$% GID %>% unique)
gidsInPhenos %<>% 
      mutate(NewName=GID) %>%
      filter(NewName %in% rownames(K)) %>% 
      bind_rows(
            gidsInPhenos %>% 
                  mutate(NewName=GID) %>%
                  filter(!NewName %in% rownames(K)) %>% 
                  left_join(
                        tibble(FullSampleName=grep("C1b",rownames(K),ignore.case = T,value = T)) %>% 
                              separate(FullSampleName,c("NewName","GBS_ID"),":",extra = 'merge',remove = F) %>% 
                              mutate(NewName=paste0("NR16",NewName)) %>% 
                              dplyr::select(-GBS_ID)) %>% 
                  filter(!is.na(FullSampleName)) %>% 
                  mutate(NewName=FullSampleName) %>% 
                  dplyr::select(-FullSampleName)) %>% 
      bind_rows(
            gidsInPhenos %>% 
                  mutate(NewName=GID) %>%
                  filter(!NewName %in% rownames(K),
                         grepl("C2a",NewName,ignore.case = T)) %>% 
                  left_join(
                        tibble(dartName=grep("C2a",rownames(K),ignore.case = T,value = T)) %>% 
                              separate(dartName,c("dartID","NewName"),"\\.\\.\\.",extra = 'merge',remove = F) %>% 
                              filter(!is.na(NewName))) %>% 
                  mutate(NewName=ifelse(!is.na(dartName),dartName,NewName)) %>% 
                  dplyr::select(-dartName,-dartID)) %>% 
      filter(NewName %in% rownames(K))
blups %<>% 
      unnest(blups) %>% 
      left_join(gidsInPhenos) %>% 
      filter(!is.na(NewName)) %>% 
      mutate(GID=NewName) %>% 
      dplyr::select(-NewName) %>% 
      group_by(Trait,OutliersRemoved) %>% 
      nest(.key = "trainingData")
blups %<>% 
      unnest(trainingData) %>% 
      bind_rows(
            iitablups %>% 
                  filter(Trait %in% blups$Trait) %>% 
                  unnest(trainingData)) %>% 
      group_by(Trait,OutliersRemoved) %>% 
      nest(.key = "trainingData") %>% 
      mutate(Dataset="IITAaugmented") %>% 
      bind_rows(blups %>% 
                      mutate(Dataset="NRCRIonly"))
rm(blupsNoOut,gidsInPhenos,iitaGIDsInPhenos,iitablups, proctime); gc()

```

## Get GEBVs
```{r, eval=F}
require(sommer)
require(furrr)
options(future.globals.maxSize= 1500*1024^2)
options(mc.cores=40)
plan(multiprocess)
proctime<-proc.time()
blups %>% 
      mutate(GEBV=future_pmap(.,function(Trait,Dataset,OutliersRemoved,trainingData){
            trndata<-trainingData %>% 
                  mutate(GID=factor(GID,levels=rownames(K)))
            fit <- mmer(fixed = drgBLUP ~1,
                        random = ~vs(GID,Gu=K),
                        weights = WT,
                        data=trndata)
            x<-fit$U$`u:GID`$drgBLUP 
            gebvs<-tibble(GID=names(x),
                          GEBV=as.numeric(x))
            saveRDS(gebvs,file=paste0("output/",
                                      "GEBV_",Trait,"_",Dataset,"_OutliersEqual",OutliersRemoved,"_NRCRI_73119.rds"))
            rm(fit); gc()}))
proc.time() - proctime

```

## Format GEBV
```{r, eval=F}
library(tidyverse); library(magrittr);
gebvs<-list.files() %>% 
      grep("GEBV_",.,value = T) %>% 
      grep("_NRCRI_73119.rds",.,value = T) %>% 
      tibble(Files=.) %>% 
      mutate(Trait=gsub("GEBV_","",Files),
             Trait=gsub("_NRCRI_73119.rds","",Trait)) %>% 
      separate(Trait,c("Trait","Dataset","OutliersRemoved"),"_") %>% 
      mutate(OutliersRemoved=gsub("OutliersEqual","",OutliersRemoved),
             Files=map(Files,~readRDS(.))) %>% 
      unnest(Files)
gebvs %<>% 
      mutate(GeneticGroup=NA,
             GeneticGroup=ifelse(grepl("NR17F",GID,ignore.case = T),"C2a",
                                 ifelse(grepl("C2bF",GID,ignore.case = T),"C2b",
                                        ifelse(grepl("^F",GID) & !grepl("C1b",GID),"C1a",
                                               ifelse(grepl("C1b",GID,
                                                            ignore.case = T),"C1b","nrTP_plus_IITA"))))) 
gebvs %>% 
      count(GeneticGroup)
```

## Write GEBVs

```{r, eval=F}
gebvs %>% 
  spread(Trait,GEBV) %>% 
  group_by(Dataset,OutliersRemoved) %>% 
  nest() %>% 
  mutate(data=pmap(.,function(data,Dataset,OutliersRemoved){
    write.csv(data,file=paste0("output/GEBV_NRCRI_",Dataset,"_OutliersRemoved",OutliersRemoved,"_73119.csv"),
              row.names = F)}))
```

# Next step

[Get Genetic Gain Estimate](GetGainEst.html)
